% !TEX root = ./tech-specification.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{交易处理}
%\section{Transaction Processing}
\label{sec:tx_processing}

\name 实现了与以太坊 \cite{ETH_yellow} 相同的虚拟机。
%\name implements the same virtual machine as Ethereum \cite{ETH_yellow}. 
一个交易的执行定义了转换函数 $\Upsilon(\st,\tx,\tau)$，其与以太坊的全局状态转换函数类似。
%The execution of a transaction defines the transform function $\Upsilon(\st,\tx,\tau)$, which is similar with Ethereum's state transition function.
在下文中我们将阐述 \name 有关交易执行的具体设计。
%In what follows we focus on the \name specific designs in the execution.

\subsection{燃料，支付和抵押}
%\subsection{Gas, Payment and Collateral}
\label{subsec:gas_and_pay}

如同第 \ref{sec:tx} 节中的定义样，每笔交易 $\tx$ 的两个域 {\bf 燃料量} 和 {\bf 燃料价格} 标明交易所需燃料 $\tx_g$ 和燃料单位价格 $\tx_p$ 的具体数值。
%As defined in Section~\ref{sec:tx} every transaction $\tx$ has two fields of {\bf gasLimit} and {\bf gasPrice} that declare the specific amount of associated gas $\tx_g$ and the price $\tx_p$ of per unit gas.
开始执行一笔交易 $\tx$ 时，应支付的燃料总价为 $\tx_g \times \tx_p$；当负责支付燃料消耗费用的一方负担不起该费用时，交易 $\tx$ 将被视为无效。
%When starting the execution of a transaction $\tx$, the purchase of gas happens at the price $\tx_g \times \tx_p$ and the transaction $\tx$ is considered invalid if the actor responsible for the cost of gas consumption cannot afford such a purchase.
正常情况下，$\tx$ 的发送方 $\sender{\tx}$ 将承担燃料消耗费用。
%Normally $\sender{\tx}$, the sender of $\tx$, is responsible for the cost of gas consumption. 
当交易 $\tx$ 调用一个有燃料赞助的智能合约 $\contract$，且 $\tx$ （如第 \ref{sec:sponsor} 节中所写明的）有资格使用赞助，
%In case the transaction $\tx$ is calling a smart contract $\contract$ with sponsorship for gas consumption and $\tx$ is qualified for the subsidy as specified in Section~\ref{sec:sponsor}, 
若 $\contract$ 有足够的 \textbf{燃料资助余额}，$\contract$ 将负责支付燃料费用；否则，交易发送方 $\sender{\tx}$ 仍将承担所有燃料消耗费用。
%$\contract$ is responsible for the purchase of gas if it has sufficient \textbf{sponsor balance for gas},and otherwise the sender $\sender{\tx}$ is still responsible for the whole purchase of gas as if there were no sponsor at all.
与以太坊相似，燃料不存在于交易执行以外。
%Like in Ethereum, gas does not exist outside the execution of transactions.


未被使用的燃料将在交易 $\tx$ 执行后退回，但退回额度不超过购买燃料所花总价的四分之一。
%The unused gas can be refunded after the transaction $\tx$ is executed, but no more than a quarter of the total value spent on purchasing. 
因此，\emph{可退还燃料量} $g^{\dagger}$ 等于以下两个中更小的值：$\tx$ \textbf{燃料量} 的四分之一和 \emph{合理剩余燃料} $g'$。即，$g^{\dagger}\eqdef \min\set{g', \tx_g/4}$。
%Thus, the \emph{refundable amount of gas} $g^{\dagger}$ is the minimum of the \emph{legitimately remaining gas} $g'$ and a quarter of the \textbf{gasLimit} of $\tx$,i.e. $g^{\dagger}\eqdef \min\set{g', \tx_g/4}$, 
特别地，若发送方导致 $\tx$ 执行失败，则理论上没有燃料可退还（即 $g^{\dagger} = g' = 0$）。
%where in principle no gas is refundable (i.e. $g^{\dagger} = g' = 0$) if the execution of $\tx$ fails due to the sender's fault. 
最初为 $\tx$ 购买燃料的一方将收到退款 $g^{\dagger}\times\tx_p$。而支付给已消耗燃料的 \coinsign 为
%The actor who initially purchased the gas for $\tx$ will get the refund of $g^{\dagger}\times\tx_p$. And the \coinsign paid for the consumed gas is 
	%
	\begin{align}
		\left(\tx_g-g^{\dagger}\right)\times \tx_p.
	\end{align}
	%
如果 $\tx$ 未被执行（即，如第 \ref{sec:tx validate} 中定义一般，仅当 $R_z=2$ 时），没有燃料需要被支付。
	%If the $\tx$ is not executed (i.e. only when $R_z=2$ as in Section~\ref{sec:tx validate})), no gas will be charged. 
若发送方 $\sender{\tx}$ 无法负担燃料费，发送方的全部余额将被用于支付燃料费，此实际收取的燃料费将被记录在收据 $R_f$ 中。
	%If the sender $\sender{\tx}$ can not afford gas fee, all its the remained balance of sender $\sender{\tx}$ will be charged as gas fee. The actual charged gas fee will be record in receipt $R_f$. 


支付完毕的燃料费将加入矿工奖励池。因此，更高的燃料价格使发送者花费更多的同时，也将增加该交易被及时处理的概率。
%The charged gas fee is added to the reward pool for miners. Thus in general a higher gas price on a transaction would cost the sender more but also increase the chance of being processed timely.


计算整个区块的累计燃料消耗时，不可退还的燃料 $g^{\dagger} - g'$ 将不被纳入考量。
%In computing the accumulated gas used of the whole block, the non-refundable gas $g^{\dagger} - g'$ is not taken into consideration. 
但当发送方 $\sender{\tx}$ 无法负担燃料费时，即使实际收取的燃料费少于 $\tx_p\times\tx_g$，我们也认为消耗的燃料量即是 $\tx_g$。
%But in case the sender $\sender{\tx}$ can not afford gas fee, the gas used is considered as $\tx_g$, even if the actual charged gas fee is less than $\tx_p\times\tx_g$. The gas used is also recorded in receipt $R_g$. 
因此，一个区块的累计燃料消耗本质上等于交易收据里 $R_g$ 域的总和。
%Thus the accumulated gas used of a block is intrinsically taking the summation of field $R_g$ over transaction receipts.


每笔交易都有一个 \textbf{存储上限} 域，表明当前交易可增加的最大存储字节 $\tx_\ell$。
%Every transactions also have a filed of \textbf{storage limit} that declare the maximum storage bytes $\tx_\ell$ increasing for the present transaction. 
执行交易前，除燃料费用和转账额以外，发送方 $\sender{\tx}$ 必须有足够余额为所需存储上限支付存储抵押，即 $\tx_\ell\times\collateralperbyte\;\unit$。
%Before transaction execution, besides gas fee and transferred value, the sender $\sender{\tx}$ must have enough balance for storage collateral for specified storage limit, i.e., $\tx_\ell\times\collateralperbyte\;\unit$. 
不同于燃料费，在执行交易前这些抵押将不被收取或锁定。
%Unlike the gas fee, these collateral will not be charged or locked at this time. 
在交易执行完成时，若发送方没有足够余额为实际增加的存储支付抵押，或实际增加的存储字节超过了存储上限，交易执行将失败。
%At the end of transaction execution, if the sender doesn't have enough balance paying for the increased storage collateral or the increased storage bytes of sender exceeds storage limit, the transaction execution fails. 
更多有关存储抵押的细节详见第 \ref{sec:collateral} 节。
%More details for collateral for storage is specified in section~\ref{sec:collateral}. 

\subsection{交易前验证}
%\subsection{Pre-execution Validation}
\label{sec:tx validate}

在被执行前，一笔处于待执行列中的交易 $\tx$ 必须通过以下有关交易合法性的第二轮测试：
%Before being executed, a transaction $\tx$ in the processing queue must pass the following secondary test of intrinsic validity. 
\begin{enumerate}[nosep]
	\item 当前纪元处于 \textbf{epochHeight} 所指明的区间之内，即当前纪元高度在 $[\tx_e - \txepochbound, \tx_e + \txepochbound]$ 之间。
	%\item The current epoch is in the range specified by \textbf{epochHeight}, i.e. current epoch height is in $[\tx_e - \txepochbound, \tx_e + \txepochbound]$.
	
	\item 交易 \textbf{序列号} 合法，即 $\tx_n = \st\left[\sender{\tx}\right]_n$，其中 $\st$ 表示当前世界状态。
	%\item The transaction \textbf{nonce} is valid, i.e. $\tx_n = \st\left[\sender{\tx}\right]_n$ where $\st$ is the current world-state.

	\item 接收方地址合法，即 $\tx_a$ 的（前 $4$ 位）类型标志属于 $\set{\typereserved,\typenormal,\typecontract}$。
    %\item The recipient address is valid , i.e. the type indicator (first $4$-bit) of $\tx_a$ belongs to $\set{\typereserved,\typenormal,\typecontract}$.
\end{enumerate}

需要注意的是交易的本地合法性，即 $\rlp$ 格式和签名有效性，
%Note that the local legality of the transaction, e.g. the $\rlp$ format and the validity of signature, 
已在有关交易合法性的第一轮测试中（定义见第 \ref{sec:block validate} 节）得到验证。第一轮测试在将相应区块纳入 \name 树图时进行，因此在第二轮测试中本地合法性将不被二次检查。
%is already verified in the first intrinsic validity test before accepting the corresponding block into the \name \tg, as discussed in Section~\ref{sec:block validate}, and will not be checked again at this moment.

若 $\tx$ 未能通过这些检查，系统将不执行该交易，不增加账户序列号，也不对此交易收取交易费用。
%If $\tx$ fails at these checks, the transaction will not be executed, the nonce for account will not increase and no transaction fee is charged for such transaction. 
用 $R'$ 表示前一笔交易的收据，
%Let $R'$ be the receipt of last transaction.
则当前交易的收据将被设为：
%Then the receipt of current transaction will be set as follows:
\begin{align}
	R_u=R'_u && R_f=0 && R_g=0 && R_{\bf l}=\emptystring && R_z=2 && R_s=0 && R_{\bf o}=\emptystring && R_{\bf i}=\emptystring
\end{align}
%
Log $R_{\bf l}$ 的布隆过滤器 $R_b$ computed accordingly。
%(The bloom filter $R_b$ of log $R_{\bf l}$ is computed accordingly. )

若 $\tx$ 通过了以上交易前检查，$\tx$ 的执行过程如本节余下内容所述。
If $\tx$ passes all the above pre-execution checks, the execution of $\tx$ is as specified in the rest of this section.

\subsection{Transaction Execution}

\subsubsection{Eligibility for sponsorship}

When $\tx$ is calling a contract $\contract$, the gas fee and storage collateral during the transaction execution could be sponsored by the contract $\contract$. If the sender $\sender{\tx}$ is in the {\bf sponsor whitelist} of contract $\contract$, the gas fee of the current transaction does not exceed the {\bf sponsor limit for gas} of contract and contract $\contract$ has a {\bf sponsor for gas}, transaction $\tx$ is eligible for sponsorship on gas consumption. Formally, we define 
\begin{align}
	\mathsf{GasElig}(\st,\tx)\eqdef \quad \mathrm{Type}_{\tx_a}=\typecontract \;\wedge\; \mathsf{Whitelist}(\st,\sender{\tx},\contract) \;\wedge\; \st[\tx_a]_p[\mathsf{gas}]_a\neq 0 \;\wedge\; \tx_g\times\tx_p \le \st[\tx_a]_p[\mathsf{limit}] 
\end{align}
where function $\mathsf{Whitelist}(\cdot)$ is defined in \cref{eq:whitelist}. 

If the sender $\sender{\tx}$ is in the {\bf sponsor whitelist} of contract $\contract$ and contract $\contract$ has a {\bf sponsor for collateral}, transaction $\tx$ is also eligible for sponsorship on storage collateral.
\begin{align}
	\mathsf{ColElig}(\st,\tx)\eqdef \quad \mathrm{Type}_{\tx_a}=\typecontract \;\wedge\; \mathsf{Whitelist}(\st,\sender{\tx},\contract) \;\wedge\; \st[\tx_a]_p[\mathsf{col}]_a\neq 0
\end{align}

\subsubsection{Preprocessing}
\label{subsubsec:preprocessing}

In the preprocessing phase of $\tx$, the balance of $\sender{\tx}$ (and the sponsor, if applicable) is examined so that the payment for any further operation is assured.
The world-state will be transformed from $\st$ into $\st^0\eqdef \st^{**}$ if $\tx$ passes the preprocessing, or directly into $\st'$ and the execution is aborted if $\tx$ fails at any step.

\paragraph{Nonce incremental.}
The beginning of execution 
causes an irrevocable changed to the state $\st$: 
the nonce of the sender, $\sender{\tx}_n$, is incremented by one. 
%
We define the state $\st^*$:
\begin{align}
	\st^*  &\eqdef \st \qquad \mbox{  except:}\\
	\st^*\left[\sender{\tx} \right]_n &\eqdef \st\left[\sender{\tx} \right]_n+1 
\end{align}

\paragraph{Gas consumption payment validation.}

The up-front payment of a transaction $\tx$ first figures out whether the gas consumption is sponsored. $\tx$ is sponsored on gas consumption if $\tx$ is eligible for sponsorship on gas consumption and the calling contract has sufficient \textbf{sponsor balance for gas fee}. 
\begin{align}
	\mathsf{GasSpr}(\st,\tx) \eqdef\quad  \mathsf{GasElig}(\st,\tx) \;\wedge\; \st[\tx_a]_p[{\sf gas}]_b\ge \tx_g\times\tx_p
\end{align}
\begin{itemize}
	\item If the gas consumption of $\tx$ is sponsored, the world-state $\st^{**}$ after gas consumption payment is as follows: 
	\begin{align}
		\st^{**}  &\eqdef \st^* \qquad \mbox{  except:}\\
		\st^{**}\left[\contract_{addr}\right]_p[{\sf gas}]_b &\eqdef \st^*\left[I_a\right]_p[{\sf gas}]_b-\tx_g\times\tx_p
	\end{align} 
	
	\item Otherwise, the sender $\sender{\tx}$ is required to pay for the gas consumption. 
	The balance of $\sender{\tx}$ should satisfy $\st^{*} \left[\sender{\tx}\right]_b \ge \tx_g\times\tx_p+\tx_v$ and otherwise a \emph{not enough balance exception} is generated. The handling of \emph{not enough balance exception} will be discussed later.
	The world-state after the gas consumption payment is defined as: 
	\begin{align}
		\st^{**}  &\eqdef \st^* \qquad \mbox{  except:}\\
		\st^{**} \left[\sender{\tx}\right]_b &\eqdef \max\set{\st^*\left[\sender{\tx}\right]_b-\tx_g\times\tx_p,0}
	\end{align}
\end{itemize}

\paragraph{Storage limit validation.}

After charging, Conflux decides who is responsible for storage collateral. If $\tx$ is eligible for sponsorship on storage collateral and calling contract $\contract=\tx_a$ has enough \textbf{sponsor balance for collateral}, contract $\contract$ is responsible for the storage collateral resulted in the execution of $\tx$ and will be the owner of modified entries. 
%
Otherwise, the sender $\sender{\tx}$ is the owner of modified entries
and has the obligation to pay corresponding storage collateral. 
%
\begin{align}
	\mathsf{ColSpr}(\st,\tx) &\eqdef\quad  \mathsf{ColElig}(\st,\tx) \;\wedge\; \st[\tx_a]_p[{\sf col}]_b\ge \tx_\ell\times 10^{18}/1024 \\
	\mathsf{ColOwner}(\st,\tx) &\eqdef\left\{ \begin{array}{ll}
		\tx_a & \text{if } \mathsf{ColSpr}(\st,\tx) = \true \\ 
		\sender{\tx} & \text{if } \mathsf{ColSpr}(\st,\tx) = \false \\ 
	\end{array}\right.
\end{align}

If $\sender{\tx}$ is the storage owner but his balance cannot afford the full collateral as declared in {\bf storageLimit} after transferring value $\tx_v$, 
i.e. $\st^{**}[\sender{\tx}]_b<\tx_v+\tx_\ell\times 10^{18}/1024$, 
then the execution of $\tx$ fails due to \emph{not enough balance exception}. 

\paragraph{Handling not enough balance exception.} 

Whenever the preprocessing of $\tx$ generates a \emph{not enough balance exception} during preprocessing, the execution of $\tx$ fails and there will be no further execution of $\tx$. To figure out whether this exception caused by the insufficient sponsorship balance in contract, the sender balance before transaction execution (i.e. $\st[\sender{\tx}]_b$) is compared with a \emph{minimum required balance} defined as
\begin{align}
	\tx_v+(1-\mathbb{I}(\mathsf{GasElig}(\st,\tx)))\times \tx_g \times \tx_p + (1-\mathbb{I}(\mathsf{ColElig}(\st,\tx)))\times \tx_\ell \times \collateralperbyte.
\end{align}

If $\st[\sender{\tx}]_b$ has enough balance for \emph{minimum required balance}, the sender $\sender{\tx}$ is considered not responsible for the generated \emph{not enough balance exception}. 
In this case, the resultant world-state $\st'$ is reverted to $\st$, the nonce of sender is reset so that $\tx$ is reusable. The receipt is composed as follows (where $R'$ refers the receipt of last transaction):
\begin{align}
	&R_u=R'_u && R_f=0 && R_g=0 && R_{\bf l}=\emptystring \\
	&R_z=2 && R_s=0 && R_{\bf o}=\emptystring && R_{\bf i}=\emptystring
\end{align}

In other cases, sender $\sender{\tx}$ is responsible for the exception. The resultant world-state is $\st'\eqdef \st^{**}$ and the receipt is composed as follows (where $R'$ refers the receipt of last transaction):
\begin{align}
	&R_u=R'_u+\tx_g && R_f=\min\{\tx_g\times\tx_p,\st[\sender{\tx}]_b\} && R_g=\mathsf{GasElig}(\st,\tx) && R_{\bf l}=\emptystring \\
	&R_z=1 && R_s=0 && R_{\bf o}=\emptystring && R_{\bf i}=\emptystring
\end{align}


\subsubsection{Execution Substate}
\label{subsubsec:substate}

The \emph{transaction substate} $A$ is a five tuple which accrues intermediate information during execution. 
\begin{align}
	A\eqdef \left( A_{\bf s}, A_{\bf l}, A_{\bf t}\right)
\end{align}
The components of $A$ are defined as follows: 
\begin{itemize}[nosep]
	\item $A_{\bf s}$ is the self-destruct set of accounts that will be discarded upon the transaction's completion.

	\item $A_{\bf l}$ is the log series consisting of indexable ``checkpoints'' in the VM code execution, allowing light clients to track the execution of a contract.

	\item $A_{\bf t}$ is the set of touched accounts, of which the empty ones will be deleted on the transaction's completion.
\end{itemize}

The empty substate $A^0$, which is also the initial substate, has no self-destructs, no logs, no touched accounts, and zero refund. Formally, $A^0$ is defined as
\begin{align}
	A^0\eqdef \left( \varnothing, \emptystring, \varnothing\right)
\end{align}

For any two substate $A^1$ and $A^2$, the accrued substate $A\eqdef A^1\Cup A^2$ is defined by 
\begin{align}
	A_{\bf s} &\eqdef A^1_{\bf s} \cup A^2_{\bf s} \\ 
	A_{\bf l} &\eqdef A^1_{\bf l} \cdot A^2_{\bf l} \\
	A_{\bf t} &\eqdef A^1_{\bf t} \cup A^2_{\bf t} 
\end{align}


\subsubsection{Type dependent execution}

If transaction passes the preprocessing, 
then {\name} evaluates the \emph{post-execution provisional state} $\st^P$ from \emph{pre-execution provisional state} $\st^0$ depending on the transaction type as specified in $\tx_a$: either contract creation or message call. 
%
The gas available for the proceeding computation is $g\eqdef \tx_g-g_0$, where $g_0$ is the intrinsic cost of $\tx$ as in (\ref{def:g0}). 

We define the tuple of post-execution provisional state $\st^{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
\begin{align}\label{def:transform}
	(\st^{P},g',A,z)\eqdef
	\begin{cases}
		\creation(\st^0,\sender{\tx},\sender{\tx},\emptystring, \mathsf{ColOwner}(\st,\tx),g,\tx_p,\tx_v,\tx_{\bf i},0,\top) &  \tx_a=\varnothing \\
		\execution(\st^0,\sender{\tx},\sender{\tx},\tx_a,\emptystring,\mathsf{ColOwner}(\st,\tx),\tx_a,g,\tx_p,\tx_v,\tx_v,\tx_{\bf d},0,\top) & \tx_a\neq\varnothing
	\end{cases}
\end{align}
%
Notice that we have three more parameters compared with Ethereum. 

The specifications of function $\creation$ and $\execution$ are given in Section~\ref{sec:creation} and Section~\ref{sec:execution} respectively.

\subsubsection{Postprocessing}\label{sec:tx_post_process}

\paragraph{Storage collateral refund and charge.}

After the message call or contract creation is processed, Conflux checks whether the incremental storage exceeds storage limit specified in $\tx_\ell$ and if the storage owner has enough balance for storage collateral. 
Let $i\eqdef \mathsf{ColOwner}(\st,\tx)$ be the address who owns modified storage entries and $v$ be the available balance to pay for storage collateral, which is defined as 
\begin{align}
	v \eqdef \begin{cases}
		\st^{P}[\sender{\tx}]_b & \text{if }\mathsf{ColSpr}(\st,\tx)=\false \\
		\st^{P}[\tx_a]_p[{\sf col}]_b &  \text{if }\mathsf{ColSpr}(\st,\tx)=\true
	\end{cases}
\end{align}
%
Notice that $\st^P[i]_o-\st^0[i]_o$ is the incremental storage collateral during execution.
If $\st^P[i]_o-\st^0[i]_o>\min\{v,\tx_\ell\times 10^{18}/1024\}$, then the execution fails because of exceeding the storage limit, 
and all the modified state will be reverted to $\st^0$, 
i.e. $\st'\eqdef\st^0$. 
Let $R'$ denote the receipt of last transaction.
Then the receipt of current transaction $\tx$ will be 
\begin{align}
	&R_u=R'_u+\tx_g && R_f = \tx_g\times\tx_p && R_g = \mathsf{GasSpr}(\st,\tx) && R_{\bf l}=\emptystring \\
	&R_z=1 && R_s = \mathsf{ColSpr}(\st,\tx) && R_{\bf o}=\emptystring && R_{\bf i}=\emptystring
\end{align}

Otherwise \name charges and refunds storage collateral and transforms world-state $\st^P$ into $\st^*$. 
We skim the self-destructed contracts here because their storage collateral have been refunded during self-destruction. 
The storage collateral in account state is also updated at this time. 

\begin{align}
	&\st^1  \eqdef \st^{P} \qquad \mbox{  except:}\\
	&\forall a \in \B_{160} \text{ with } \st^{P}[a]_o\neq\st^0[a]_o, \\
	&\quad \begin{cases}
	\st^1[a]_p[{\sf col}]_b \eqdef \st^{P}[a]_p[{\sf col}]_b - (\st^{P}[a]_o-\st^0[a]_o) & \mbox{if $a$ refers to a contract account, i.e. $\mathsf{Type}_{a}=\typecontract$} \\
	\st^1[a]_b \eqdef \st^{P}[a]_b - (\st^{P}[a]_o-\st^0[a]_o) & \mbox{if $a$ refers to a normal account, i.e. $\mathsf{Type}_{a}= \typenormal$}
	\end{cases}\\
	&\st^1[a_{\sf stake}]_{\bf s}[k_4] \eqdef \st^P[a_{\sf stake}]_{\bf s}[k_4] + \sum_{a\in \B_{160}}(\st^{P}[a]_o-\st^0[a]_o) \\
	&\mbox{where:}  \\
	&a_{\sf stake} \eqdef \stakingcontract \\ 
	&k_4 \eqdef {\sf [total\char`_storage\char`_tokens]_{\sf ch}} 
\end{align}

\paragraph{Sponsor balance for collateral refund.}

After refunding collateral for storage, the sponsor balance for collateral of all the destructed contract are refunded to their sponsor.

\begin{align}
	&\st^2  \eqdef \st^1 \qquad \mbox{  except:}\\
	&\forall a \in \B_{160},\;  \st^2[a]_b\eqdef \st^1[a]_b + \sum_{a'\in A_{\bf s}} \mathbb{I}(\st^1[a']_p[{\sf col}]_a=a)\times \st^1[a']_p[{\sf col}]_b \\
	&\forall a' \in A_{\bf s},\; \st^2[a']_p[{\sf col}]_b\eqdef 0
\end{align}


\paragraph{Gas fee refund.}

The \emph{refundable amount of gas} $g^{\dagger}$ is the minimum of the \emph{legitimately remaining gas} $g'$ (as calculated in (\ref{def:transform})) and a quarter of the \textbf{gasLimit} of $\tx$,
	i.e. $g^{\dagger}\eqdef \min\set{g', \tx_g/4}$.
The refund of gas fee is applied on world-state $\st^{*}$ and results in $\st'\eqdef \Upsilon(\st,\tx)$.

\begin{align}
	& \st'  \eqdef \st^2 \qquad \mbox{  except:}\\
	& \quad \begin{cases} 
		\st'\left[\tx_a\right]_p[{\sf gas}]_b \eqdef \st^2\left[\tx_a\right]_p[{\sf gas}]_b+g^{\dagger}\times \tx_p 
		& \mbox{if $\mathsf{GasSpr}(\st,\tx)=\true$}\\
		\st' \left[\sender{\tx}\right]_b \eqdef \st^2\left[\sender{\tx}\right]_b + g^{\dagger}\times \tx_p 
		& \mbox{if $\mathsf{GasSpr}(\st,\tx)=\false$}
	\end{cases} 
\end{align}

\paragraph{Transaction Receipt.} 

Now the transaction execution is accomplished.
The returning status code $z$ denotes whether the execution succeeds or not. 
Supposing that $R'$ is the receipt of last transaction, 
the receipt of current transaction will be as follows:
\begin{align}
	\begin{array}{llll}
		R_u=R'_u+g' & R_f = (\tx_g-g^{\dagger})\times \tx_p & R_g = \mathsf{GasSpr}(\st,\tx) & R_{\bf l}=A_{\bf l} \\ 
		R_z=z & \multicolumn{3}{l}{R_s = \left\{\begin{array}{ll}
			\mathsf{ColSpr}(\st,\tx) & \text{if }z=0\\
			0 & \text{if }z=1
		\end{array}\right.} \\
		\multicolumn{4}{l}{R_{\bf o}=\mathsf{ToList}\left(\left\{ (a,\st^P[a]_o - \st^0[a]_o) | a\in \B_{160}\;\wedge\; \st^P[a]_o - \st^0[a]_o>0 \right\}\right)}\\
		\multicolumn{4}{l}{R_{\bf i}=\mathsf{ToList}\left(\left\{(a,\st^0[a]_o - \st^P[a]_o) | a\in \B_{160}\;\wedge\; \st^P[a]_o - \st^0[a]_o<0  \right\}\right)}
	\end{array}
\end{align}
